---
title: "Análisis y curación de datos"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Grupo 2

> Integrandes:

* Martín Hunziker
* Claudio Sarate
* Ramiro Caro

## Resolución Ejercicio 1

### Descripción del dataset

Todos los vuelos que partieron de NYC (JFK, LGA or EWR) en el año 2013

Columna           | Descripción
------------------| -------------
year,month,day    | Date of departure 
dep_time,arr_time | Actual departure and arrival times (format HHMM or HMM), local tz.
sched_dep_time,sched_arr_time | Scheduled departure and arrival times (format HHMM or HMM), local tz.
dep_delay,arr_delay | Departure and arrival delays, in minutes. Negative times represent early departures/arrivals.
hour,minute | Time of scheduled departure broken into hour and minutes.
carrier | Two letter carrier abbreviation. See airlines() to get name
tailnum | Plane tail number
flight  | Flight number
origin,dest | Origin and destination. See airports() for additional metadata.
air_time | Amount of time spent in the air, in minutes
distance | Distance between airports, in miles
time_hour |Scheduled date and hour of the flight as a POSIXct date. Along with origin, can be used to join flights data to weather data

### Importación y estadisticas del dataset


```{r summary}
library(nycflights13)
fligths <- nycflights13::flights
flights
colnames(flights)
summary(flights)
```


#### 1. Listar los vuelos que arribaron con un retraso de mas de dos horas.

```{r Pregunta 1}
fly_delay <- subset(fligths, fligths$dep_delay>120)
dim(fly_delay)
fly_delay
```

#### 2. Listar los vuelos que volaron hacia Houston (IAH o HOU)
```{r Pregunta 2}
fly_Houston <- subset(fligths, (fligths$dest=="HOU" | fligths$dest=="IAH"))
dim(fly_Houston)
fly_Houston
```

#### 3.Listar los vuelos que fueron operados por United, American o Delta.
```{r Pregunta 3}
fly_Am_carrier <- subset(fligths, (fligths$carrier=="UA" | fligths$carrier=="AA" | fligths$carrier=="DL"))
dim(fly_Am_carrier)
fly_Am_carrier

```

#### 4. Listar los vuelos que salieron en Verano (Julio, Agosto y Septiembre)

```{r Pregunta 4}
fly_verano <- subset(fligths, (fligths$month>= 7 & fligths$month <=9 ))
dim(fly_verano)
fly_verano

```
#### 5.Listar los vuelos que arrivaron mas de dos horas tarde, pero salieron bien.

Se considera que los vuelos que salieron adelantados o en horario (dep_delay = 0) como vuelos que salieron bien

```{r Pregunta 5}
fly_DepOK_ArDelay <- subset(fligths, (fligths$dep_delay<= 0 & fligths$arr_delay >=120 ))
dim(fly_DepOK_ArDelay)
fly_DepOK_ArDelay

```

#### 6.Listar los vuelos que a salieron entre medianoche y las 6 am.

```{r Pregunta 6}
fly_nocturno <- subset(fligths, (fligths$dep_time>= 0 & fligths$dep_time<= 600 ))
dim(fly_nocturno)
fly_nocturno

```
